### 2017-10-30 tex/article.tex

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Journal Article
% LaTeX Template
% Version 1.4 (15/5/16)
%
% This template has been downloaded from:
% http://www.LaTeXTemplates.com
%
% Original author:
% Frits Wenneker (http://www.howtotex.com) with extensive modifications by
% Vel (vel@LaTeXTemplates.com)
%
% License:
% CC BY-NC-SA 3.0 (http://creativecommons.org/licenses/by-nc-sa/3.0/)
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%----------------------------------------------------------------------------------------
%	PACKAGES AND OTHER DOCUMENT CONFIGURATIONS
%----------------------------------------------------------------------------------------
%for images

% /\/\/\/\/\/\/\/\/\/\/\/\/\/\/\/\/\/\/\/\/\/\/\/\/\/\/\/\/\/\/\/\/\/\/\/\/\/\/\
%
% THIS IS IMPORTANT! RUN WITH pdflatex -shell-escape TO AVOID EPS CONVERSION ERROR
%
% /\/\/\/\/\/\/\/\/\/\/\/\/\/\/\/\/\/\/\/\/\/\/\/\/\/\/\/\/\/\/\/\/\/\/\/\/\/\/\

\documentclass[twoside,twocolumn]{article}

\usepackage{graphicx}

\graphicspath{ {../img/} }

\usepackage{subfig}
\usepackage{todonotes}
\usepackage[sc]{mathpazo} % Use the Palatino font
\usepackage[T1]{fontenc} % Use 8-bit encoding that has 256 glyphs
\linespread{1.05} % Line spacing - Palatino needs more space between lines
\usepackage{microtype} % Slightly tweak font spacing for aesthetics
\usepackage[english]{babel} % Language hyphenation and typographical rules
\usepackage[hmarginratio=1:1,top=32mm,columnsep=20pt]{geometry} % Document margins
\usepackage[labelfont={bf,up},textfont={it,up}]{caption} % Custom captions under/above floats in tables or figures
\usepackage{booktabs} % Horizontal rules in tables
\usepackage{lettrine} % The lettrine is the first enlarged letter at the beginning of the text

\usepackage{enumitem} % Customized lists
\setlist[itemize]{noitemsep} % Make itemize lists more compact

\usepackage{abstract} % Allows abstract customization
\renewcommand{\abstractnamefont}{\normalfont\bfseries} % Set the "Abstract" text to bold
\renewcommand{\abstracttextfont}{\normalfont\small\itshape} % Set the abstract itself to small italic text

\usepackage{titlesec} % Allows customization of titles
\renewcommand\thesection{\Roman{section}} % Roman numerals for the sections
\renewcommand\thesubsection{\roman{subsection}} % roman numerals for subsections
\titleformat{\section}[block]{\large\scshape\centering}{\thesection.}{1em}{} % Change the look of the section titles
\titleformat{\subsection}[block]{\large}{\thesubsection.}{1em}{} % Change the look of the section titles

\usepackage{fancyhdr} % Headers and footers
\pagestyle{fancy} % All pages have headers and footers
\fancyhead{} % Blank out the default header
\fancyfoot{} % Blank out the default footer
\fancyhead[C]{Bird simulation $\bullet$ \today $\bullet$ v1} % Custom header text
\fancyfoot[RO,LE]{\thepage} % Custom footer text

\usepackage{titling} % Customizing the title section
\usepackage{hyperref} % For hyperlinks in the PDF
\usepackage{amssymb,amsmath,bbold,braket}

%----------------------------------------------------------------------------------------
%	TITLE SECTION
%----------------------------------------------------------------------------------------

\setlength{\droptitle}{-4\baselineskip} % Move the title up

\pretitle{\begin{center}\Huge\bfseries} % Article title formatting
\posttitle{\end{center}} % Article title closing formatting
\title{Simulating a flock of birds} %Simulating a flock of birds
\author{%
\textsc{Steffen Randrup, Michel Smola, Andrei Timus}\\[1ex] % Your name
\normalsize University of Copenhagen \\ % Your institution
%\normalsize \href{mailto:john@smith.com}{john@smith.com} % Your email address
%\and % Uncomment if 2 authors are required, duplicate these 4 lines if more
%\textsc{Jane Smith}\thanks{Corresponding author} \\[1ex] % Second author's name
%\normalsize University of Utah \\ % Second author's institution
%\normalsize \href{mailto:jane@smith.com}{jane@smith.com} % Second author's email address
}
\date{\today} % Leave empty to omit a date
\renewcommand{\maketitlehookd}{%
\begin{abstract}
\noindent
% In this paper we present a model for studying the behaviour of a group of particles which have some level of interaction between them. The model was inspired from the~\cite{Vicsek} paper the particles have a constant absolute velocity and at each time step  assume the average direction of nearby neighbours with some added perturbation, $\eta$.
%The behaviour was studied by analysing how the average velocity of the flock $v_a$ varied when parameters like noise, density, number of particles where changed. In addition to $[1]$ we have introduced a cone of vision for every particle and studied the behaviour of the flock in the same conditions. From the simulation results it has been observed that the flock goes through a phase transition where all the particles assume the same direction. The average velocity $v_a$ played the role of an order parameter and it scaled with $(\eta_c-\eta)^\beta$, where $\beta$ is a critical exponent.
The motion of a flock of birds appears complex and difficult to classify due to
the large number of paticipants each with their own idea of where to go.
Yet such flocks display clear collective motion. In this paper we describe such a flock
similar to a system of self-propelled particles interacting with nearby neighbours.
We extend this model by reducing the interaction of any given bird by only interacting with other birds
within the bird's vision. Any bird will act as an average of any other birds it can see.
We show that this simple model of limited interaction may cause the flock to align
over distances larger than the extend of a single bird's interaction. Including how
the noise in each bird's direction and density of birds affect the alignment.
We compare the transition from unordered to ordered motion to a phase transition and
comment on the effects of a vision based approach to this transition.
\end{abstract}
}

%----------------------------------------------------------------------------------------

\begin{document}
\maketitle

%----------------------------------------------------------------------------------------
%	ARTICLE CONTENTS
%----------------------------------------------------------------------------------------

\section{Collective motion of birds}

One may consider the collective movement of birds to be described by a simple set of rules.
In this paper we describe the movement of such a flock as a system of self-propelled particles.
Our model is an extension of the model presented by Tamás Vicsek and his group in their paper ``Novel Type of Phase Transition in a System of Self-Driven Particles''~\cite{Vicsek}.
The motion of each bird or particle is governed by the motion of other birds or particles nearby.
In the following we may use birds and particles interchangably.

In our setup the particles where distributed randomly in a square lattice of size $L$.
Each particle had assigned a random direction $\theta$ and a constant interaction radius $r$.
One update consisted in modifying the position of all particles according to the following equation:

\begin{equation}
x_{i}(t+1)=x_i(t)+v_i(t)\Delta t
\end{equation}

where $\Delta$t represents the time between 2 updates.
The velocity of a particle had a constant absolute value and a direction given by:

\begin{equation}
\theta(t+1)=\braket{\theta(t)}_r+\Delta \theta
\end{equation}

where

\begin{equation}
  \braket{\theta(t)}_r=\arctan\frac{\braket{\sin(\theta(t))}_r}{\braket{\cos(\theta(t))}_r}
\end{equation}
%
is the average direction of the velocities of nearby particles wich are inside the radius of influence, $r$, of the given particle.
$\Delta \theta$ represents the noise in the system and takes values with equal probability within the interval $[-\eta/2,\eta/2]$.
For a constant number of particles there are 3 parameters which can be varied: $\eta$,$\rho$,and $v$,
where $v$ is the distance a particle travels between 2 updates.
$\eta$ describes the interval in which $\Delta\theta$ is picked. $\rho$ is the density of the particles.
This is effectively a way of controlling the size of the lattice if the number of particles is fixed.

We further extend the model by implementing a cone-like vision. This means, that 
rather than averaging the velocities over all birds within a certian radius, we 
only look at a slice of this circle. For a bird to be within the vision of 
another bird, it must be within the interaction radius, $r$, and the angle 
between the bird's velocity and direction to the other bird must be within $\pm \theta_{\text{cone}}$. The 
limit $\theta_{\text{cone}} = \pi$ equals the case without limiting the
cone vision and reproduces the results from~\cite{Vicsek}.

\begin{figure}[!htb]
	\centering
	\subfloat{{\includegraphics[width=\columnwidth/2]{low-eta-low-rho}}}
  \subfloat{{\includegraphics[width=\columnwidth/2]{low-eta-high-rho}}}
	\qquad
	\subfloat{{\includegraphics[width=\columnwidth/2]{high-eta-low-rho}}}
	\subfloat{{\includegraphics[width=\columnwidth/2]{high-eta-high-rho}}}
	\caption{Different configurations of $\eta$ and $\rho$\\Top left: low $\eta$ low $\rho$\\Top right: Low $\eta$, high $\rho$\\Bottm left: High $\eta$, low $\rho$\\Bottom right: High $\eta$, high $\rho$}\label{fig:configs}
\end{figure}

Figure~\ref{fig:configs} shows the configuration of birds for different values of $\rho$ and $\eta$.  
For low values of both $\rho$ and $\eta$ the birds tend to form smaller groups in which the birds move in the same direction.
The groups themselves move randomly.
When the randomness is high and the density low, the overall motion of the birds is just random.
In the case of high randomness and high density, the birds appear to move randomly but with an overall trend in directions
shared among the all the birds.
If the density is high enough, while the noise has low values, the birds move orderly in one direction.
This overall trend in motion is similar to the system going through a kinetic phase transition
where all the particles share the same direction. In order to characterize the behaviour of the flock
we have determined the average normalized velocity of the system and investigated how it changes when the input parameters are modified:

\begin{equation}
v_a=\frac{1}{Nv}\left\vert\sum_{i=1}^{N} v_i\right\vert
\end{equation}

For disordered motion $v_a$ is aproximatively 0, while for ordered $v_a$ has the value near 1.

The behaviour of $v_a$ is found to be similar to that of the order parameter of some equilibrium systems close to their critical point. For large system sizes the data shows scaling which is an indication of a phase transition. Equations~\eqref{eq:va-eta} and~\eqref{eq:va-rho} may be used to describe such a system near the critical value, at which the phase transition occours.

\subsection{Determining the critical noise level}
To determine the critical noise parameter $\eta_c$ we used a quadratic fit model
to to fit the falling curve for $\eta < \eta_c$ and a linear fit for $\eta > 
\eta_c$, both visible in figure~\ref{fig:va_over_eta}. $\eta_c$ is then set to 
the intersection of the fit curves. Hereby we determine
\begin{align}
\eta_c = 4.56***
\end{align}
\todo[inline]{put in final number}

\begin{equation}
  \label{eq:va-eta}
  v_a \approx{[\eta_c(\rho)-\eta]}^\beta 
\end{equation}
\begin{equation}
  \label{eq:va-rho}
  v_a \approx{\left[\rho-\rho_c (\eta)\right]}^\gamma
\end{equation}

% \section{Results}
\section{From disorder to order}

The alignment, $v_a$, depends on the amount of randomness $\eta$ in the system 
and the density $\rho$, as seen in equations~\eqref{eq:va-eta} and~\eqref{eq:va-rho}.
In figure~\ref{fig:va_over_eta} we see how the alignment of 
the birds decreases as the randomness of their motion increases. The decrease 
is to be expected, as a higher degree of randomness should disrupt alignment. 
This reflects to $v_a$, which approaches $0$ in a totally unordered state and $1$
in a totally ordered state.
What might be more interesting is the general shape of the curves. We may then 
determine $\beta$ from equation~\eqref{eq:va-eta} to be \todo[inline]{get value of $\beta$}
This is also shown in figure~\ref{fig:criticaleta}. The results presented here 
are all computed without the conevision (equivalent to $\theta_{\text{cone}} = \pi$). 
Figure~\ref{fig:va_over_eta} shows how $v_a$ varies with the noise $\eta$ and 
density $\rho$ for a constant value of density and noise respectively.
As seen in the graph the overall alignment of the particles depends 
on the noise and for a high number of particles N the system becomes less ordered. 
For high $\eta$ the flock reaches a coherent moving phase, while for low $\eta$ 
the particles move randomly in the lattice.

\begin{figure}[!htb]
  \centering
  \includegraphics[width=\columnwidth]{va_over_eta}
  \caption{Dependence of alignment on randomness}\label{fig:va_over_eta}
\end{figure}

\begin{figure}[!htb]
	\centering
	\includegraphics[width=\columnwidth]{va_over_etac_minus_eta_over_etac}
	\caption{Determining $\beta$}\label{fig:criticaleta}
\end{figure}

Similarly we may see the effects of the density on the alignment of birds. This also fits the prediction from equation~\eqref{eq:va-rho}. This is shown in figure~\ref{fig:va_over_rho}.

\begin{figure}[!htb]
	\centering
	\includegraphics[width=\columnwidth]{va_over_rho}
	\caption{Dependance of alignment on density}\label{fig:va_over_rho}
\end{figure}

\subsection*{Cone-like vision}\label{subsec:conevision}

\todo[inline]{The results in this section were based on code producing wrong results, thus it must be rewritten}

% The effects of our implementation of an angle based vision are only detectable on both low randomness and low density.
% This is expected as a higher density causes an effective long range interaction, which may span the entire flock.
% This negates the effect of an angle based vision as the flock moves in a coherent fashion.
% As for randomness of motion it decreases the overall alignment of the flock as seen in figure~\ref{fig:va_over_eta}.
% In figure~\ref{fig:conevision} we present the overall alignment's dependance of the angle describing this cone-like
% vision for different densities. The randomness is kept at a low value of $\eta=0.1$.
% The results correspond quite well with our expectation of increasing the alignment with larger angle.
% For an angle of $\pi$ we see, the alignment looks like it did when we considered every bird within the interaction radius.
% We also notice the effects of the density of birds. As described above we expect a long range interaction
% between the birds to arise when the density increases.
% This effect is visible in the figure as we see the alignment increase faster, when the density is higher.
% Again we see, how the system goes between an unordered state and an ordered state. For low densities we don't reach full alignment, but that might be due to the lack of interaction between far-away birds. The trend however is similar to the phase transition presented earlier.

The cone-like vision has no or little effect on the final alignment of the birds, only the time it takes to reach this level of alignment.
What we do see however is the shape of the formed groups. In stead of forming dense groups, the birds now form lines or open
triangles whose shapes are defined by the angle of vision. An example of such a configuration is seen in Figure~\ref{fig:angle-birds}.
In Figure~\ref{fig:conevision} we see how the effect of the cone-like vision has little effect on the final alignment. There appears to be a slight upward trend for the density $\rho=0.1$. This may be attributed to the time taken to reach alignment as each data point has been taken after the same amount of time. It may not have been long enough for the smaller angles. For the cases $\rho=1,0.01$ They both reach the alignment shown in Figure~\ref{fig:va_over_rho}. The runtime before measuring $v_a$ has for Figure~\ref{fig:conevision} been twice as long as for the other figures.

\begin{figure}[!htb]
\begin{center}
\includegraphics[width=\columnwidth]{va_over_angle}
\end{center}
\caption{Conevision}\label{fig:conevision}
\end{figure}

\begin{figure}[!htb]
\begin{center}
\includegraphics[width=\columnwidth]{angles}
\end{center}
\caption{Angles  ..... }\label{fig:angle-birds}
\end{figure}

\section{Discussion}

We have compared the collective motion of birds to a phase transition.
The connection between these is not obvious. However we do see simlilarities in how increasing the noise in the system
will distrupt alignment and how denser systems will display longer effective interaction range. Causing the entire
system to behave coherently.

The addition of a cone-like vision is, as the name might suggests, inspired from birds rather than the study of phase transitions.
\todo[inline]{The previous discussion of cone-like vision was wrong, so rewrite it}
%The effects of the vision based approach were suttle. They were only detectable for low densities and noise, and for
%sufficiently large angles of vision the system would again display behavior to that without the vision based approach.
%As discussed this behavior is similar to that of a phase transition, however it is only detectable for low denisities.
%This also means, that the system doesn't align completely in the range where it is detectable,
%unlike for the other transitions presented in this paper.
We have treated the angle of vision more like a property
of the system than a variable. This is a naive approach as it assumes that any single bird only ever looks in one direction.

The values of $v_a$ have been taken as a time average after some transient time after which the system has settled into a state, which doesn't change significantly.

\section{Conclusion}

% In this paper we have presetend a model for studying the behaviour of interacting particles. The interaction between particles consisted in maintaining the average velocity of nearby neighbours. The system was studied when particles shared a common interaction radius and when each particle had a cone of vision. During the simulation it has been observed that the system goes through a phase transition where all particles assume the same direction. In order to describe the transition we have calculated the average normalized velocity va and investigated how it behaves when we modify the parameters of the system. In the thermodynamic limit va can be described using (5) or (6).
% By plotting these equations we have found  the critical exponent $\beta$ to be equal to\todo[inline]{$\beta$}

In this paper we have presented a model for studying the motion of a flock of birds.
The model contained contained few variables; density of birds, noise in their direction and an angle
controlling which other birds influence a given bird's motion. Each of these variables gave rise
to the movement of the flock going from disordered to ordered. For the case of the angle of vision
the transition was only detectable for low densities and low noise as higher densities made more
birds be within sight and high noise reduced alignment enough for the effect of the vision to be hard to detect.

%----------------------------------------------------------------------------------------
%	REFERENCE LIST
%----------------------------------------------------------------------------------------

\begin{thebibliography}{99} % Bibliography - this is intentionally simple in this template

  \bibitem{Vicsek}Tamás Vicsek, András Czirók, Eshel Ben-Jacob, Inon Cohen and Ofer Shochet\\
  Novel Type of Phase Transition in a System of Self-Driven Particles\\
  Physical Review Letters, Volume 75, Number 6, August 1995

\end{thebibliography}

%----------------------------------------------------------------------------------------

\end{document}